Essays in Computational Management Science

### 2018-05-11 Chapters/hathor-bitcoin-math.tex
Hathor: finished \o/

@@

It is easy to integrate $u(t)$ because $\frac{1+abx}{1+bx} = \frac{1-a}{1+bx} + a \Rightarrow \int_0^t u(x) dx = at + \frac{1-a}{b} \log(1+bt)$. So,
$$F_T(t) = 1 - (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$
$$f_T(t) = \lambda \left( \frac{1+abt}{1+bt} \right) (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$

@@

Thus,
\begin{align*}
\mathcal{L}\{F_T(t)\} &= \mathcal{L}\{1 - (1+bt)^n e^{-\lambda at}\} \\
	&= \mathcal{L}\{1\} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \tag{$\mathcal{L}$ is a linear operator} \\
	&= \frac{1}{s} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \\
	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \mathcal{L}\{t^k e^{-\lambda at}\} \\
	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \frac{k!}{(s+\lambda a)^{k+1}}
\end{align*}

@@

Let's check this equation for already known scenarios. When $a=1$, then $n=0$ and $\mathbf{E}[T] = 1/\lambda$. When $b \rightarrow +\infty$, it reduces to the case in which the hash rate is multiplied by $a$, which we have already studied. In fact, $n \rightarrow 0$, $u(t) \rightarrow a$, and $\mathbf{E}[T] = \frac{1}{\lambda a}$.

@@

For instance, if we would like to know the impact of smoothly double the hash rate in the next week, then the parameters would be $\lambda = 1/600$, $a=2$, $M=1\text{ week}=3600\cdot24\cdot7 = 604,800$, $b$ can be calculated using $\epsilon = \frac{a-1}{1+bM} < 0.01 \Leftarrow b > 0.000163690 \Leftarrow n < 10.1818$. So, for $n=10$, then $b=0.000166666$ and $\epsilon = 0.009823 < 0.01$, as expected. Finally, $\mathbf{E}[T] = 557.65$. In other words, during the next week, the average time between blocks will be 9 minutes and 17 seconds, instead of the normal 10 minutes. If the hash rate had suddenly doubled, the average time between blocks would be 5 minutes.

@@

The function $\mathbf{E}(T)|_{t_1}^{t_2}$ may be used to a piecewise linear analysis of any hash rate change. Let's analyze the hash doubling in one week. Then, $u(1 \text{week}) = u(604800) = 2$, thus $a^2 = \frac{1}{604800} \Rightarrow a = \frac{1}{120 \sqrt{42}}$. Let's sample the interval $[0, 1 \text{week}]$ every hour, i.e., ($t_0$, $t_1$, $t_2$, \dots, $t_168$), where $t_i = i \cdot 1 \text{hour} = i \cdot 3600$.

For each point $t_i$, let $g_i(t) = [H(t-t_i) - H(t-604800)] u(t) + 2 H(t-604800)$, then $\mathbf{E}(T) = \mathbf{E}(T)|_{t_i}^{t_{i+1}} + 1 - e^{\frac{t}{\eta}}$. The result is presented in Figure .

@@

\begin{figure}[ht]
\centering\includegraphics[width=\textwidth]{./images01/bitcoin-H-increasing.png}
\caption{$\mathbf{E}[T]$ when H increases linearly with $u(t) = 1 + at$.\label{fig-bitcoin-H-increase}}
\end{figure}

@@

\citet{karame2012two}

There are many possible ways to attack the Bitcoin (REF). In this section, we are interested in a particular attack: the double spending attack.

@@

Analyze when $k=2$ (for instance, Binance requires only two confirmations for deposits with less than $1,000$).

@@

Let's assume that all transactions have exactly the same size $S$ and pay exactly the same fee. If the Bitcoin block's maximum size is $M$, there would be room for $s = \lfloor M/S \rfloor$ transactions in each block. Using the results from \citet{bailey1954queueing}, we have found that $\pi_n = \frac{z_s-1}{z_s^{n+1}}$ is the probability of having $n$ unconfirmed transactions in the pool subjected to $s > m$, where $m=\frac{\lambda_{\text{TX}}}{\lambda_{\text{blocks}}}$ and $z_s$ is the single root of the polynomial $z^s(1+m(1-z))-1$ with $|z_s|>1$. The average size of the unconfirmed transaction pool is $\mathbf{E}(\pi) = \frac{1}{z_s - 1}$. Thus, $s > m \Rightarrow \lambda_{\text{TX}} < s \lambda_{\text{blocks}}$. The probabilities form a simple geometric series, and so they are exponentially decreasing. We may interpret it as a stable system, i.e., the unconfirmed transactions pool size is finite. In the Bitcoin's network, $\lambda_{\text{blocks}} = 1/\eta$ and the average number of transactions per block is $s=2,250$, so, this solution is valid when $\lambda_{\text{TX}} < 2,250/600 = 3.75$. Hence, $3.75$ is the maximum number of new transactions per second that the Bitcoin's network may handle. When $\lambda_{\text{TX}} > 3.75$, the unconfirmed transaction pool starts to grow indefinitely.

The average waiting time of a transaction to be confirmed, when $m < s$, is $\mathbf{E}(w) = \frac{1}{\lambda_{\text{TX}} (z_s-1)}$.

When $m \ll s$, $z_s \rightarrow 1 + 1/m$. So, the average number of unconfirmed transactions is $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = \eta = 600 \text{ seconds}$. In the Bitcoin's network, $s \gg m \Rightarrow \lambda_{\text{TX}} \ll 3.75$.

When $\lambda_{\text{TX}} \rightarrow s$, $z_s \rightarrow 1$. So, $\mathbf{E}(\pi) \rightarrow +\infty$.

Finally, we conclude that the Bitcoin's network capacity is $\lambda_{\text{blocks}} s = s/\eta = s/600$ transactions per second, where $s$ is the average number of transactions per block.

>>>>

\subsection{Hash rate suddenly changing}

@@

It is easy to integrate $u(t)$ because $\frac{1+abx}{1+bx} = \frac{1-a}{1+bx} + a$, which yields $\int_0^t u(x) dx = at + \frac{1-a}{b} \log(1+bt)$. So,
$$F_T(t) = 1 - (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$
$$f_T(t) = \lambda \left( \frac{1+abt}{1+bt} \right) (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$

@@

Let $\mathcal{L}$ be the Laplace Transform. Thus,
\begin{align*}
\mathcal{L}\{F_T(t)\} &= \mathcal{L}\{1 - (1+bt)^n e^{-\lambda at}\} \\
	&= \mathcal{L}\{1\} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \tag{$\mathcal{L}$ is a linear operator} \\
	&= \frac{1}{s} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \\
	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \mathcal{L}\{t^k e^{-\lambda at}\} \\
	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \frac{k!}{(s+\lambda a)^{k+1}}
\end{align*}

@@

Let's check this equation for already known scenarios. When $a=1$, then $n=0$ and $\mathbf{E}[T] = 1/\lambda$. When $b \rightarrow +\infty$, it reduces to the case in which the hash rate is multiplied by $a$, which we have already studied. In fact, $b \rightarrow +\infty$ yields $n \rightarrow 0$, $u(t) \rightarrow a$, and $\mathbf{E}[T] = \frac{1}{\lambda a}$.

@@

For instance, if we would like to know the impact of smoothly double the hash rate in the next week, then the parameters would be $\lambda = 1/600$, $a=2$, $M=1\text{ week}=3600\cdot24\cdot7 = 604,800$, $b$ can be calculated using $\epsilon = \frac{a-1}{1+bM} < 0.01$, which yields $b > 0.000163690$ and $n < 10.1818$. So, for $n=10$, then $b=0.000166666$ and $\epsilon = 0.009823 < 0.01$, as expected. Finally, $\mathbf{E}[T] = 557.65$. In other words, during the next week, the average time between blocks will be 9 minutes and 17 seconds, instead of the normal 10 minutes. If the hash rate had suddenly doubled, the average time between blocks would be 5 minutes.

@@

\subsection{Comparison of the models}

In order to compare the hash rate change models, namely (i) suddenly change, (ii) smoothly change, and (iii) linear change, I have applied each of them to the same scenarios. In the first scenario, the hashrate will double in the next week, whereas, in the second scenario, it will halve in the next week.

In both the smoothly change model and the linear change model, I could have calculated the average time between blocks during the whole week. It would not give us much information, because the time between models would be increasing (or reducing) more and more as the days goes by. And we are really interested in the average time between blocks throughout the days, and not the average of the whole week.

Thus, I have analyzed a piecewise hash rate change, i.e., I have calculated the average time between blocks for each hour throughout the week. First, I split the whole week into $24 \cdot 7$ intervals, $(t_0, t_1, t_2, \dots, t_168)$, where $t_i = 3600i$. Then, I calculated the average for each interval $(t_k, t_{k+1})$. Let $H_k^0$ and $H_k^1$ be the initial and final hash rate of the $(t_k, t_{k+1})$ interval. So, I also ensured the continuity of the hash rate between consecutive intervals, i.e., $H_k^1 = H_{k+1}^0$.

I compared both the smoothly change model and the linear change model with the suddenly change model. The difference between them is negligible. Let $\epsilon$ be the maximum absolute error between the models, than $\epsilon < 0.8$ and $\epsilon/H < 0.2\%$, for all intervals. The maximum absolute error between the linear and the suddenly change models can be seen in Figure \ref{fig:bitcoin-H-change-error}.

\begin{figure}[!htb]
\centering \includegraphics[width=0.7\textwidth]{./images01/H-max-abs-error.png}

\caption{Maximum absolute error between the linear and the suddently change models. \label{fig:bitcoin-H-change-error}}
\end{figure}

Therefore, we may conclude that it is reasonable to approximate the average time between blocks using only the suddenly change model in each interval of one hour.

The average time between blocks throughout the days can be seen in Figure \ref{fig:bitcoin-H-change}. It was calculated using the suddenly change model with the hash rate changing linearly during the week.

\begin{figure}[!htb]
\centering
\subfloat[Doubling the hash rate]{\includegraphics[width=\textwidth]{./images01/H-double-week.pdf}}

\subfloat[Halving the hash rate]{\includegraphics[width=\textwidth]{./images01/H-halves-week.pdf}}

\caption{The average time between blocks when the hash rate changes over time. \label{fig:bitcoin-H-change}}
\end{figure}

@@

%\begin{figure}[ht]
%\centering\includegraphics[width=\textwidth]{./images01/bitcoin-H-increasing.png}
%\caption{$\mathbf{E}[T]$ when H increases linearly with $u(t) = 1 + at$.\label{fig-bitcoin-H-increase}}
%\end{figure}

@@

There are many possible ways to attack the Bitcoin \citep{karame2012two, heilman2015eclipse, bahack2013theoretical, bonneau2016buy, neudecker2015simulation}. In this section, we are interested in a particular attack: the double spending attack.

@@

% TODO Analyze when $k=2$ (for instance, Binance requires only two confirmations for deposits with less than $1,000$).

@@

Let's assume that all transactions have exactly the same size $S$ and pay exactly the same fee. If the Bitcoin block's maximum size is $M$, there would be room for $s = \lfloor M/S \rfloor$ transactions in each block.

Using the results from \citet{bailey1954queueing}, we have found that $\pi_n = \frac{z_s-1}{z_s^{n+1}}$ is the probability of having $n$ unconfirmed transactions in the pool subjected to $s > m$, where $m=\frac{\lambda_{\text{TX}}}{\lambda_{\text{blocks}}}$ and $z_s$ is the single root of the polynomial $z^s(1+m(1-z))-1$ with $|z_s|>1$. In this case, the average size of the unconfirmed transaction pool is $\mathbf{E}(\pi) = \frac{1}{z_s - 1}$.

When $s > m$, the probabilities $\pi_n$ form a simple geometric series with common ratio smaller than one, which means the probabilities are exponentially decreasing. Since $\pi_n \rightarrow 0$ when $n \rightarrow \infty$, we may interpret it as a stable system, i.e., the unconfirmed transactions pool size is finite.

When $s \le m$, the system is unstable, which means the unconfirmed transactions pool size keeps growing towards infinity. In this case, the system is not capable of processing the demand for a long period of time.

Using the fact that $m=\frac{\lambda_{\text{TX}}}{\lambda_{\text{blocks}}}$ and $\lambda_{\text{blocks}} = 1/\eta$, the stability condition $s > m$ is reached when $\lambda_{\text{TX}} < s/\eta$.

In the Bitcoin's network, the average number of transactions per block is $s=2,250$, so, the system is stable when $\lambda_{\text{TX}} < 2,250/600 = 3.75$ tx/s. Therefore, $3.75$ is the maximum number of new transactions per second that the Bitcoin's network may handle. When $\lambda_{\text{TX}} > 3.75$ tx/s, the unconfirmed transaction pool starts to grow indefinitely.

When the system is stable, the average waiting time of a transaction to be confirmed is $\mathbf{E}(w) = \frac{1}{\lambda_{\text{TX}} (z_s-1)}$.

$m \ll s$ yields $z_s \rightarrow 1 + 1/m$. Thus, the average number of unconfirmed transactions $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = \eta = 600 \text{ seconds}$. In the Bitcoin's network, $m \ll s$ is reached when $\lambda_{\text{TX}} \ll 3.75$ tx/s. In other words, when the number of new transactions per second is way smaller than 3.75 tx/s, the average waiting time of a transaction to be confirmed is 600 seconds, which means, on average, all transactions will be confirmed in the next block.

But, $\lambda_{\text{TX}} \rightarrow s/\eta$ yields $z_s \rightarrow 1$. Hence, $\mathbf{E}(\pi) \rightarrow +\infty$, which means the system is going towards instability.

Therefore, we conclude that the Bitcoin's network capacity is $\lambda_{\text{blocks}} = s/\eta = s/600$ transactions per second, where $s$ is the average number of transactions per block.

For instance, in order to be a stable system and process 15 transactions per second, each block would have to confirm, on average, 9,000 transactions. Bitcoin's network is really far from this point.
